Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-10 examples.tex
+

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} < 1 \land (\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;return;
;;..
;;clean up the stack
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed. Not here }
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !! \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$

@@

\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).}

>>>>

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{ \forall_{i\in \textsf{0 ... 511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
\subseteq
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;clean up the stack
..
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed. Not here }
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
\subseteq
$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{  \textsf{entry\_val + KERNBASE}_{\textsf{pfn}}  \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val}  \textsf{entry\_val}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$

@@

\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates). Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$ }
